---
title: "lab_02"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

### **Q1** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most?

```{r}
#string detect and filter#
Climate <- umd_courses |>
  filter(str_detect(title, "Climate")) |>
  filter(seats >= 1)
    
Climate |> 
  group_by(department) |> 
  summarise(
    count_department = n()) |>
arrange(desc(count_department))

```

**A1: 26 departments offer courses with the word "Climate" and at least one seat. Atmospheric and Oceanic Science has offered the most.**

------------------------------------------------------------------------

### **Q2** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top departments different in each result? Where is Journalism in those results?

```{r}
#greater than 0#
# fall 2023 term = 202308

umd_courses |>
  filter(term == 202308) |>
  group_by(department) |>
  summarise(
    count_classes = n()
  ) |>
 arrange(desc(count_classes))
  
umd_courses |>
  filter(syllabus_count > 0) |> filter(term == 202308) |>
  group_by(department) |>
  summarise(
    count_department = n()
  ) |>
 arrange(desc(count_department))
  

```

**A2: The departments that have the most courses in the Fall 2023 term are Business and Management (164 classes) and School of Music (137 classes). Journalism offered the 6th highest number of courses in fall 2023 (110). 125 departments have courses with at least one syllabus posted. Business and Management (82 classes) and Economics (48 classes) have the highest number of courses with syllabi posted. Journalism ranks 38th in number of courses with syllabi posted (12).**

------------------------------------------------------------------------

### **Q3** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.

```{r}
#lubridate#
maryland_expenses |>
  mutate(expenditure_date=mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date, "month")) |>
  group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```

**A3: It looks like spending increased in 2022, compared to the previous year. The top four most expensive months were from 2022, and most of the months with the least amount of spending are from 2021. It also looks like the later months of the year tend to have higher expenditures than earlier months.**

------------------------------------------------------------------------

### **Q4** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record indicated money spent in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe).

```{r}
#case when + str detect#
spending_location <- maryland_expenses |>
  mutate(
    state = case_when(
        str_detect(address, " Maryland ") ~ "In-state",
         str_detect(address, " California ") ~ "Out of state",
        str_detect(address, " Washington ") ~ "Out of state",
        str_detect(address, " Louisiana ") ~ "Out of state",
        str_detect(address, " Florida ") ~ "Out of state",
        str_detect(address, " North Carolina ") ~ "Out of state",
        str_detect(address, " Massachusetts ") ~ "Out of state",
        str_detect(address, " West Virginia ") ~ "Out of state",
        str_detect(address, " Virginia ") ~ "Out of state",
        .default = "NA"
      )
  )

spending_location |>
 group_by(state) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```

**A4: The results show that in-state spending accounts for** **\$77,723,146. Out of state spending adds up to \$33,164,994.**

------------------------------------------------------------------------

### **Q5** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.

**Which campaign employee (or payout recipient) is making the highest amount of money?**

```{r}
Salary <- maryland_expenses |> filter(expense_category == "Salaries and Other compensation")

head(Salary)

Salary |>
  group_by(payee_name) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```

**A5: It looks like the highest amount of expenses for salaries/other compensation go through payroll organizations like Paychex, Gusto, and ADP. The fourth-highest payout went to Martin-Lauer Associates, which is an organization that helps with campaign fundraising services. The highest ranked individual receiving campaign expenses was Edward Miller (number 12). I wasn't able to find any organizations or firms called Edward Miller online, so I am assuming this is an individual. It could be Edward B. Miller, Gov. Bob Erlich's Deputy Chief of Staff from 2004-2007? \$152,000 is an odd amount of money to receive from Maryland campaigns for someone who had already gotten in trouble for working alongside corrupt lobbyists. It's probably a different Edward Miller, though.**
