---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

dc <- read.csv("~/JOUR 772/Burke-JOUR-772-Assignments/major_assignments/reverse_engineering/data/dc-wikia-data.csv") |> clean_names()

marvel <- read.csv("~/JOUR 772/Burke-JOUR-772-Assignments/major_assignments/reverse_engineering/data/marvel-wikia-data.csv") |> clean_names()

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Among the characters for which we have gender data, females made up only 29.3 percent of the DC character list and 24.7 percent of the Marvel roster. 
-   **Analysis summary**: It looks like we found a slightly lower percentage of women for both DC and Marvel in our analysis. While these findings were close enough to the original findings (23.4% vs 24.7% for Marvel, and 28.5% vs. 29.3% for DC), the findings varied by about 1 percentage point. There were a number of cells in the sex category left blank: it is possible that the researches counted some of these as female in their analysis without updating the dataset, which contributed to the discrepancy between our analyses.

```{r}
# Put code to reverse engineer sentence here

dc |> group_by(sex) |> summarize(amount_women = n()) |> 
  mutate(pct_women = (amount_women/6896)*100)

marvel |> group_by(sex) |> summarize(amount_women = n()) |> 
  mutate(pct_women = (amount_women/16376)*100)
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
