---
title: "lab_06"
author: "derek willis"
date: "2022-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# load your libraries
library(tidyverse)
library(janitor)
library(lubridate)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll use some of our 911 overdose call data to do this. The first thing we want to do is to combine multiple counties' data into a single dataframe so that we can ask some questions. First, let's combine data from Cecil, Carroll and Allegany counties into a new dataframe.

*Before* you combine them you'll need to clean up some of the column names to make the data easier to work with - make sure you have loaded the library to do that. You also need to make sure that each column has the same name and datatype (you can check the datatypes using `glimpse`). If any of them does not, you need to fix that.

```{r}
# load and combine the call data from those three counties
allegany_county <- read_csv("~/JOUR 772/Burke-JOUR-772-Assignments/labs/lab_06/data/allegany_911.csv") |> clean_names()
baltimore_city <- read_csv("~/JOUR 772/Burke-JOUR-772-Assignments/labs/lab_06/data/baltimore_911.csv") |> clean_names() |> mutate(zip_code = as.character(zip_code))
carroll_county <- read_csv("~/JOUR 772/Burke-JOUR-772-Assignments/labs/lab_06/data/carroll_911.csv") |> clean_names()
cecil_county <- read_csv("~/JOUR 772/Burke-JOUR-772-Assignments/labs/lab_06/data/cecil_911.csv") |> clean_names() 
```

Then we'll join some demographic data with Baltimore City 911 calls and ask some questions of that. For that, you'll need to load the Baltimore City 911 data and the ZCTA demographic data, and join them based on the zip code. Remember, the zip codes in both dataframes must be the same datatype (and should be characters). You should start with your Baltimore 911 dataframe in creating your join.

```{r}
# load and join the Baltimore-related data
ZCTA <- read_csv("~/JOUR 772/Burke-JOUR-772-Assignments/labs/lab_06/data/maryland_zcta.csv") |> clean_names() |> mutate(zcta5n = as.character(zcta5n))

Baltimore_911_And_ZCTA <- baltimore_city |> left_join(ZCTA, join_by(zip_code == zcta5n)) 
```

## Answer questions

### Q1. Let's start with our data from Allegany, Cecil and Carroll counties. Write code to generate a dataframe that shows the total number of calls for each county. What's the order of counties from most calls to least?

A1. The county with the most calls is Carroll (448), then Cecil (438), and then Allegany (418).

```{r}
counties_combined <- bind_rows(list(allegany_county, cecil_county, carroll_county))

counties_combined |> group_by(county) |> summarise(
    count_county = n()
  ) |> arrange(desc(count_county))
```

### Q2. In your combined dataframe that you used in A1, add a column for the month of each call, then write code to show the total number of calls per county and month. Are there any outliers or noteworthy results? Describe the general pattern of the results.

Then calculate the most calls per county and location. Which location had the most calls - where and what is that location?

**A2. August seems to be a rough month for Cecil County and Allegany County, as the highest number of 911 calls occurred during this month for both counties.**

**Allegany County also seems to seems to dominate the top 6 highest number of 911 calls per month, despite coming in last place overall out of the three counties.**

**It seems like the most number of 911 calls comes from 314 Grove Neck Road in Earlville (in Cecil County), at 15 calls in the dataset.** **That's a 12 million dollar house with 9 bedrooms.**

```{r}
counties_combined |> mutate(month=month(date)) |> group_by(county, month) |> summarize(county_month = n()) |> arrange(desc(county_month))

counties_combined |> mutate(month=month(date)) |> group_by(location, county) |> summarize(location_amount = n()) |> arrange(desc(location_amount))
```

### Q3. Let's turn to the Baltimore City data. You're interested in exploring calls in zip codes with a high percentage of children under 18. Write code to isolate calls in zip codes where the percentage of under 18 population is at least 75% and then show the zip code and population and how many calls occurred in each of those zip codes.

Where are those zip codes in the city?

Add a column to your code that calculates the number of calls per 1,000 people for those zip codes. To calculate a per capita rate, you might find this short guide handy: [https://observablehq.com/\@palewire/per-capita-calculator](https://observablehq.com/@palewire/per-capita-calculator){.uri}.

Which zip code has the highest rate of calls per 1,000 people? Find the neighborhoods in that zip code that are listed in the data - you can use group_by or distinct to do this. What are some of those neighborhoods?

**A3. (See \# note below) While I wasn't able to calculate a per-capita rate, the zip code with the highest amount of calls is 21217, with 774 calls. This zip code includes Penn North, Sandtown-Winchester, and Upton neighborhoods. In second place is 21223, with 525. This zip code includes Carrollton Ridge, Broadway East, and New Southwest/Mount Clare neighborhoods.**

```{r}

Baltimore_911_And_ZCTA |> 
  group_by(zip_code, neighborhood) |> 
  summarize(zip_code_amount = n()) |>
  arrange(desc(zip_code_amount)) 
  #mutate(zip_code_amount = as.numeric(zip_code_amount)) |>
  #mutate(population = as.numeric(population)) |>
  mutate(calls_per_capita = (zip_code_amount/population)*1000)

# I wasn't able to get this last line of code to work. I have tried mutating all the variables (zip_code and zip_code amount) to numeric but it still gives the error "non-numeric argument to binary operator." 

#I am not sure why this would happen, given that zip_code_amount is an <int> and population is <num>, and according to Google <int> is a subsection of <num>. Not sure why it refused to treat one or the other as a number for my math problem. 

#I tried to fix it by mutating every variable to <num>, but I got the error that for my population = as.numeric(population) function, 'list' object cannot be coerced to type 'double'. Google's solution of unlist did not help either. 

#I was able to summarize zip code amount, but nothing beyond that. I tried all I could, but it wouldn't cooperate so here is what the rest of the code would theoretically look like if the last line worked:

#mutate(seventyfive_pct_under_18 = (population/pop_18_over)*100) |>
#arrange(desc(seventyfive_pct_under_18))

```

### Q4. Choose your own adventure: I want you to come up with a question you could ask of the Baltimore City data that combines both the calls and the demographic information. You must use both demographics and the overdose call data in your work.

**A4: Do older neighborhoods tend to call 911 more often than younger neighborhoods in Baltimore City?**

**In looking at both tables, it appears as though neighborhoods with a higher population of residents under 18 call at a significantly higher rate than neighborhoods with a higher population of 65 and over. Neighborhoods with a larger percentage of minors rack up hundreds of calls, or even over a thousand in one case. Neighborhoods with a higher percent of seniors, however, tend to only rack up a couple dozen, if that. I also noticed that Baltimore City leans really young. The neighborhood with the highest population of minors has a majority at 81.5%, whereas the neigborhood with highest population of seniors in Baltimore has only 23.8% over 65.**

```{r}
Baltimore_911_And_ZCTA |>
  group_by(pct_pop_65) |>
  summarize(call_amount = n()) |>
  arrange(desc(pct_pop_65))

Baltimore_911_And_ZCTA |>
  group_by(pct_under_18) |>
  summarize(call_amount = n()) |>
  arrange(desc(pct_under_18))
```
