---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    mean_seats = mean(seats),
  ) |>
  arrange(desc(mean_seats))
```

**A1: Engineering, Cooperative Education has the highest mean number of seats. I was surprised to see Agricultural and Resource Economics up so high at number 11. Math and Science courses were expected for me, but I didn't know the agriculture program was so large here.**

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}
umd_courses |>
  group_by(title) |>
  summarise(
    count_classes = n()
  )  |>
  arrange(desc(count_classes))
```

**A2: Most of the top classes here are internships, research classes, or independent study. The class in the data set that has been offered the highest number of times is "Doctoral Dissertation Research."**

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}
umd_courses |>
  group_by(instructors) |>
  summarise(
    count_classes = n()
  )  |>
  arrange(desc(count_classes))
```

**A3: The instructor that has taught the most classes in the data set is Elisa Barkley Brown. The top two slots are placeholder names (N/A and Instructor TBA).**

------------------------------------------------------------------------

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

```{r}


```

**A4: It would have made it easier to find the most listed professor if there was a "mode" (referring to the *mathematical* mode) function. So far I haven't been able to find one; the only result for mode(variable) shows the mode in terms of the *storage unit* of the variable (e.g. \<chr\> or \<int\>).**

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

```{r}
umd_courses |>
  group_by(id) |>
    summarise(
    count_classes = n(),
    total_seats = sum(seats)
  ) |>
  arrange(desc(total_seats))
```

**A5: I chose to look into whether or not upper-level courses tend to have more seats than lower-level courses. While I haven't been able to come up with a more efficient way to sort the data by level (because the course levels are attached to the department abbreviation), while scrolling through the first ten pages, I counted that only 29 out of the top 100 courses were over the 300 level. Most of the courses were at the 100 or 200 level.**

**This is fairly interesting, as it means that intro-level courses have more seats available than upper-level classes in the department. This could be because more students need to take these classes to fulfill a requirement (that is how it worked at my liberal arts undergrad school; students needed to take a math, science, and social science course regardless of major)? Or it could indicate a higher turnover rate in the math/science programs: they expect fewer students to continue their science studies beyond intro level?**
